---
title: "Interaction Between Transmission Type and MPG"
author: "KTezel"
date: "12/19/2020"
output:
  html_document: default
  pdf_document: default
---

# Executive Summary
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Motor Trend is a magazine about the automobile industry. It is interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome), particularly:

  “Is an automatic or manual transmission better for MPG”
  “Quantify the MPG difference between automatic and manual transmissions”

From the dataset, it is clear that there is a significant difference between the average mpg per automatic and manual transmission. To show the differences between average mpg per transmission type, a boxplot method is used to show the differences between median mpgs. Then, hypothesis testing, specifically, T-test is used to see whether the difference between the differences in means are significant or not. According to the T-test, differences were significant and that the mpg for automatic cars were higher than that of the cars with manual transmission.

# Libraries
```{r libraries}
library(ggplot2)
library(dplyr)
library(tidyr)
```

# Getting the Data
```{r data}
data(mtcars)
```

# Data Preprocessing
```{r preprocess}
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- factor(mtcars$am)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
```
# Exploratory Data Analysis
## Looking at data briefly
```{r head}
head(mtcars)
```

```{r dimensions}
dim(mtcars)
```

## Looking for missing data
There is no missing data.
```{r missing data}
colSums(is.na(mtcars))
```

## Plotting MPG and Automatic Transmission (1 = Yes, 2 = No)
From the plot graph, it looks like when the transmission is automatic, MPG increases.
Below is the boxplot graph. It can be clearly seen that the median MPG is higher for the cars with automatic transmission. 

```{r boxplot}
p2 <- ggplot(data = mtcars, aes(factor(am),mpg)) + geom_boxplot() + labs(x= "Transmission", y = "MPG", title = "MPG: AT vs MT")
p2
```

# Inference
## Are the differences in mean mpg between automatic and manual is significant? Let's use T-test to verify.

```{r t-test}
t.test(mpg ~ am, data = mtcars)
```
P-value from the t-test is very small which suggest that our null hypothesis is false, hence suggesting that the difference in mean mpg between automatic and manual is pretty significant.

# Regression Analysis
## Checking to see the relationship between MPG and Transmission
### Linear Regression
```{r lin_am}
fit_lm <- lm(mpg ~ am, data=mtcars)
summary(fit_lm)
```
Linear regression model predicts that when the car is automatic, it increases the mpg by approximately 7 miles. 
Intercept and the coefficients  are significant and the variables explain approximately 36% (R-squared 0.3598) of the variation in MPG which is not a good model since 35% R-squared value is not good enough.

## Other Regression Models
Below are some other models that has some higher R-squared ratio. For the purposes of this report, only the R-squared values will be extracted from the report.
```{r lm_final}
summary(lm(mpg~wt*factor(cyl),data=mtcars))$r.squared
summary(lm(mpg~wt+factor(cyl),data=mtcars))$r.squared
```

# Appendix
```{r plot_all_variables}
ggplot(mtcars, aes(x=wt, cyl, drat, y=mpg, group=am, color=am, height=3, width=3)) + geom_point() +  
scale_colour_discrete(labels=c("Automatic", "Manual")) + 
xlab("weight") + ggtitle("Scatter Plot of MPG vs. Weight by Transmission")
```

```{r subset}
mtcars_subset <- subset(mtcars, select = c(mpg, cyl, wt, drat, am))
```

```{r ggplot}
mtcars_subset %>%
    gather(-mpg, -am, -cyl, key="var", value = "value") %>%
    ggplot(aes(x = value, y = mpg, color = am, shape = factor(cyl))) +
    geom_point() +
    facet_wrap(~ var, scales = "free") +
    theme_bw()
```

## Residual Plot for the best model
```{r residual_plot}
fit_1 <- lm(mpg~wt+factor(cyl),data=mtcars)
par(mfrow = c(2,2))
plot(fit_1)
```

